---
title: "Early_Key_Takeways_TRF"
author: "Taylor Francisco"
date: "2023-10-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages}
library(tidyverse)
library(stargazer)
library(stringr)
library(ggplot2)
library(psych)
library(plm)
library(readr)
```


#Upload 2 data tables

```{r upload data}

setwd("/Users/taylorfrancisco/Downloads/")

CHL <- read_csv("UNHCR_CHL_2021_HFS_Q1Q2_v2.1.csv")

COL <- read.csv("UNHCR_COL_2020_HFS_Q4_v2.1.csv")
```

CHL is from data collected from Chile. COL is data collected from Colombia. 
Chile has 406 observations of 485 variables. Colombia has 622 observations of 586 variables. 
Our team selected a set of 45 total variables that are common throughout most of 
the surveys, which we are creating new data frames including only these variables 
to allow us to create a dataframe with all the surveys with minimal desktop space. 

#Data Cleaning (selecting variables)

##CHL

```{r}
CHL_clean<- CHL %>% select(CountryOfAsylum, nationality, ethnicity, intentionmove, TotalAdult, TotalMinor, AppliedRefugee, documentation_id, Childinschool, Childvirtualed,  Childwhynotschool_nodocs, Childwhynotschool_nomoney, Childwhynotschool_failedschool,  Childwhynotschool_childwork, Childwhynotschool_fearschool, Childwhynotschool_disease, Childwhynotschool_disability, Childwhynotschool_helphome, Childwhynotschool_noschools, Childwhynotschool_nointerest, Childwhynotschool_pregnancy, Childwhynotschool_nospot, Childwhynotschool_notransport, Childwhynotschool_discrimnation,  Childwhynotschool_discrimethnic, Childwhynotschool_finished, Childwhynotschool_recentlyarrive, Childwhynotschool_intransit, Childwhynotschool_noinfo, Childwhynotschool_toolate, Childwhynotschool_nolanguage, disabilityacc_disabilityserv, medicalReceived, RiskYes, RiskReturn_medical, RiskStay_medical, whynotMedical_nomoney, whynotMedical_noinsurance, whynotMedical_noID, whynotMedical_noinfo, whynotMedical_feararrested, whynotMedical_distance, whynotMedical_notavailable, whynotMedical_denied)

#missing variables from this dataset 
#Childwhynotschool_familynot.

```

New clean dataset has 44 variables, as expected with one variable not collected 
from this specific survey. 

###Add year as variable 

```{r}
CHL_clean<- CHL_clean %>% mutate(Year = 2021)
```


###Save Cleaned DF as CSV
In order to allow for this dataframe to be used by the rest of group, save as csv
with name format chosen by Data Analysis team. 

```{r}
write.csv(CHL_clean, file = "HFS_Chile_2021_Q1Q2", row.names = FALSE) 
```

##COL

```{r}

COL_clean<- COL %>% select(progres_CountryofAsylum, nationality, ethnicity, intentionmove, AppliedRefugee, documentation_id, Childinschool, Childvirtualed,  Childwhynotschool_nodocs, Childwhynotschool_nomoney, Childwhynotschool_failedschool,  Childwhynotschool_childwork, Childwhynotschool_fearschool, Childwhynotschool_disease, Childwhynotschool_disability, Childwhynotschool_helphome, Childwhynotschool_familynot, Childwhynotschool_noschools, Childwhynotschool_nointerest, Childwhynotschool_pregnancy, Childwhynotschool_nospot, Childwhynotschool_notransport, Childwhynotschool_discrimnation,  Childwhynotschool_discrimethnic, Childwhynotschool_finished, Childwhynotschool_recentlyarrive, Childwhynotschool_intransit, Childwhynotschool_noinfo, Childwhynotschool_toolate, Childwhynotschool_nolanguage, disabilityacc_disabilityserv, medicalReceived, RiskYes, RiskReturn_medical, RiskStay_medical, whynotMedical_nomoney, whynotMedical_noinsurance, whynotMedical_noID, whynotMedical_noinfo, whynotMedical_feararrested, whynotMedical_distance, whynotMedical_notavailable, whynotMedical_denied)


#rename column that isn't same as needed final variable
COL_clean = rename(COL_clean, CountryOfAsylum = progres_CountryofAsylum)


#missing variables from this dataset 
#TotalAdult, TotalMinor
```

New clean dataset has 43 variables, as expected with two variables not collected 
from this specific survey.

###Add year as variable 

```{r}
COL_clean<- COL_clean %>% mutate(Year = 2020)
```

###Save Cleaned DF as CSV

```{r}
write.csv(COL_clean, file = "HFS_Colombia_2020_Q4", row.names = FALSE) 

```


#Data Exploration 

Exploring the nationalities represented in these two data sets
```{r}
unique(CHL_clean$nationality)
unique(COL_clean$nationality)
```
#Graph of nationality 
```{r}
#combine dataframes
library(dplyr)
both <- bind_rows(CHL_clean, COL_clean)

#factor nationality

both$nationality.fact = as.factor(both$nationality)

class(both$nationality.fact)

nat_hist <- ggplot(both, aes(nationality.fact, fill = CountryOfAsylum))
nat_hist + geom_histogram(stat = "count",  binwidth = .5, position = "dodge") +
  labs( x = "Nationality", y = "count", 
        title ="Nationality of Respondents") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```
  From the two countries of asylum (where the refugee is currently located), the 
vast majority of the refugees are coming originally from Venezuela. This makes
sense because both of this countries are in Latin America, so it would be 
surprising to see refugees from another continent (e.g. Syria).
  I am interested to see what it means that some of the refugees in Colombia say 
their nationality is Colombia. Are they internally displaced and thus counted in
the umbrella of refugees and internally displaced persons? We will have to check
how the surveys we are using define refugee/IDPs. 
  There is also a small amount of people who did not report there nationality 
that does not have a label in this initial graph. 
 Overall main takeaway is that themajority of refugees from these two surveys 
are from Venezuela. 

#Graph of Intention to move 
```{r}
#factor intentionmove
unique(both$intentionmove)

both$intentionmove.fact = as.factor(both$intentionmove)
#both$intentionmove <- factor(both$intentionmove.fact, levels = c(1, 2, 3, 4), 
                        #labels = c("stay", "relocate", "dontknow","notell" ), exclude = NA, ordered= FALSE)

class(both$intentionmove.fact)




move_hist <- ggplot(both, aes(intentionmove.fact, fill = CountryOfAsylum))
move_hist + geom_histogram(stat = "count",  binwidth = .5, position = "dodge") +
  labs( x = "Intention to Move", y = "count", 
        title ="Intention to Move by Respondents") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

```
  In both Chile and Colombia, the majority of refugee respondents intend to stay
in the country they are seeking asylum in. There seems to be two much smaller 
groups that are roughly equal of those who do not know what they plan to do and 
those who plan to relocate. 
  There is a very small amount of respondents that did not tell the surveyors 
what their intentions are. 
  It seems like Colombia and Chile are the final destination for these refugees. 
I wonder if the other surveys my colleagues are analyzing show similar patterns
in the countries included in their data sets. 




